---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300,
  fig.align = "center"
)
```

# pedFamilias

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/pedFamilias)](https://CRAN.R-project.org/package=pedFamilias)
[![](https://cranlogs.r-pkg.org/badges/grand-total/pedFamilias?color=yellow)](https://cran.r-project.org/package=pedFamilias)
[![](https://cranlogs.r-pkg.org/badges/last-month/pedFamilias?color=yellow)](https://cran.r-project.org/package=pedFamilias)
<!-- badges: end -->

The purpose of **pedFamilias** is to facilitate data exchanging between the **pedsuite** packages and the [Familias](https://familias.name/) software. The main functions are `readFam()` and `writeFam()` for reading and writing .fam files associated with Familias. Notably, .fam files written with the DVI module of Familias are supported, and also database files without pedigree information. 

These functions were originally part of [**forrel**](https://github.com/magnusdv/forrel), but were split out to streamline maintenance and provide a more lightweight import for other packages otherwise independent of **forrel**.

## Installation
Install **pedFamilias** from CRAN as follows:
```{r eval = F}
install.packages("pedFamilias")
```

Alternatively, you may install the development version from GitHub:

```{r eval = F}
# install.packages("remotes")
remotes::install_github("magnusdv/pedFamilias")
```

## Example
```{r, message = FALSE}
library(pedFamilias)
```

For a simple illustration of **pedFamilias** we read the example file `paternity.fam` shipped with the package:

```{r paternity-load}
fam = system.file("extdata", "paternity.fam", package = "pedFamilias")

peds = readFam(fam)
```

Here are the pedigrees, including genotypes for the first marker:
```{r paternity-peds, fig.height = 2.5, fig.width = 4.5, out.width = "60%"}
plotPedList(peds, hatched = typedMembers, marker = 1)
```

Further analysis of the data may be carried out with the [**forrel**](https://github.com/magnusdv/forrel) package. For instance, the following command computes the likelihood ratio comparing the two hypotheses:

```{r}
forrel::kinshipLR(peds)
```

